
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{6.6 - Doppler Effect}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (1999)  What is a “Doppler Effect”?
\item (1999)  A whistle sound of frequency $ 1200$ Hz was directed to an approaching train moving at $ 48$ km$/$h​ . The whistle-man then listened to the beats between the emitted sound and that reflected from the train. What is the beat frequency detected by the whistle-man?
\item (2000)  Write two uses of Doppler effect.
\item (2000)  An observer standing by a railway track notices that the pitch of an engine whistle changes in the ratio of $ 5$:$ 4$ on passing him. What is the speed of the engine?
\item (2007)  What is meant by Doppler effect? 
\item (2007)  Mention two $ (2)$ common applications of the Doppler shift. 
\item (2007)  Ultra sound of frequency $ 5 \times 10^{6}$ Hz is incident at an angle of $ 30^{\circ}$ to the blood vessel of a patient and a doppler shift of $ 4.5$ KHz is observed. If the blood vessel has a diameter $ 10^{-3}m$ and the velocity of ultrasound is $ 1.5 \times 10^{3}$  $ m/s$ . Calculate the:
 \begin{itemize}
\item blood flow velocity. 
\item volume rate of blood flow. 
\end{itemize}
\item (2015)  What is meant by Doppler effect?
 \begin{itemize}
\item Write down three uses of Doppler effect.
\end{itemize}
\item (2015)  A whistle emitting a sound of frequency $ 440$ Hz is tied to a string of $ 1.5$ m length and rotated with an angular velocity of $ 20$ rad$/$s in the horizontal plane.  Calculate the range of frequencies heard by an observer stationed at a large distance.
\item (2015)  A police on duty detects a drop of a $ 10\%$ in the pitch of the horn of a motor car as it crosses him. Calculate the speed of the car.
\item (2016)  Ultrasound of frequency $ 4.0$ MHz is incident at an angle of $ 30^{\circ}$ to a blood vessel of diameter $ 1.6$ mm.  If a Doppler shift of $ 3.2$ kHz is observed, calculate the blood flow velocity and the volume rate of blood flow.  Assume that the speed of ultrasound is $ 1.5$ km$/$s.
\item (2016)  The absorption spectrum of a faint galaxy is measured and the wavelength of one of the lines identified as the calcium $ H$ line is found to be $ 478$ nm.  The same line has a wavelength of $ 397$ nm when measured in a laboratory. 
 \begin{itemize}
\item Is the galaxy moving towards or away from the observer on the Earth?
\item Determine the speed of the galaxy relative to observer on the Earth.
\end{itemize}
\item (2017)  A cyclist and a railway train are approaching each other with a speed of $ 10$ m$/$s and $ 20$ m$/$s respectively. If the engine driver sounds a warning siren at a frequency of  $ 480$ Hz, calculate the frequency of the noise heard by the cyclist:
 \begin{itemize}
\item Before the train has passed.
\item After the tram has passed. 
\end{itemize}
\item (2019)  What is Doppler effect? 
\item (2019)  The cyclist moving at $ 10$ m$/$s and the railway train at $ 20$ m$/$s are approaching each other. If the engine driver sounds a warming siren at a frequency of $ 480$ Hz:
 \begin{itemize}
\item calculate the frequency of the note heard by the cyclist before and after the train has passed away. 
\end{itemize}
\end{itemize}

\end{document}