
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{12.4 - Environmental Pollution}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (1998)  Define ionosphere.
\item (1998)  Mention the ionospheric layers that exist during the day time.
\item (2000)  What is the importance of the following layers of the atmosphere?
 \begin{itemize}
\item The lowest layer
\item The ionosphere
\end{itemize}
\item (2007)  Give a summary of location, constitution and practical uses of the stratosphere, ionosphere, and mesosphere.
\item (2010)  Define the ionosphere and give one basic use of it.
\item (2010)  Why is the ionosphere obstacle to radio astronomy?
\item (2013)  Explain why the small ozone layer on the top of the stratosphere is crucial for human survival
\item (2013)  Electrical properties of the atmosphere are significantly exhibited in the ionosphere.
 \begin{itemize}
\item  What is the layer composed of and what do you think is the origin of such constituents.
\item  Mention two uses of the ionosphere.
\end{itemize}
\item (2013)  Briefly explain on the following types of environmental pollution:
 \begin{itemize}
\item  Thermal pollution.
\item  Water pollution.
\end{itemize}
\item (2014)  Describe the sources and effects of the following pollutants on the environment:
 \begin{itemize}
\item Air pollution. 
\item Radiation pollution.
\end{itemize}
\item (2016)  What is meant by aerial environment?  Give two examples.
\item (2016)  Describe three ways at which the aerial environment is threatened.
\item (2017)  Give two factors which determine whether a planet has an atmosphere or not.
\item (2017)  Briefly explain the major causes of the following types of environmental pollution:
 \begin{itemize}
\item Water pollution. 
\item  Air pollution.
\end{itemize}

\end{itemize}

\end{document}