
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{10.2 - Semiconductors}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (1998)  Describe the function of each of;
 \begin{itemize}
\item the electron gun
\item the deflection system and
\item the display system of the Cathode ray Oscilloscope.
\end{itemize}
\item (1999)  Distinguish between insulators, semi-conductors and metals as far as conduction is concerned.
\item (2000)  Distinguish between metals and semiconductors in terms of energy bands. 
\item (2007)  How does the arrangement of the energy level in a semiconductor differ from that of an insulator?
\item (2013)  Mention one application of LED. What type of a semiconductor is it?
\item (2014)  What is light emitting diode (LED).
\item (2014)  Give three advantages of LED's lamp in radio and other electronic system over filament lamps.
\item (2014)  What is the basic difference between good conductors and semiconductors.
\item (2015)  Mention four important properties of a semiconductor.
\item (2015)  Applying the concept of doping, explain how a free electron and a positive charge can be created in a semiconductor crystal. 
\item (2016)  What is the importance of doping as applied to semiconductors?
\item (2016)  Distinguish between $ n-$ type and $ p-$ type semiconductors.  Give three points.
\item (2016)  Discuss the mode of action of each of the following sensors:
 \begin{itemize}
\item Thermistor (TH).
\item Light Dependent Resistor (LDR).
\end{itemize}
\item (2017)  Define the term semiconductor.
 \begin{itemize}
\item Give three examples of semiconductor materials. 
\end{itemize}
\item (2017)  Outline two factors on which electrical conductivity of a pure semiconductor depends. 
\item (2017)  Explain the meaning of the following terms:
 \begin{itemize}
\item $ P-$ type semiconductor.
\item $ N-$ type semiconductor. 
\end{itemize}
\item (2018)  List two chief properties of semiconductors. 
\item (2018)  Why is it easier to establish the current in a semiconductor than in an insulator?
\item (2018)  Distinguish between conductors and semiconductors on the basis of their energy band structures. 
\end{itemize}

\end{document}