
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{10.6 - Telecommunication}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (1998)  Give the reason for better reception of radio waves for high Frequency signals at night than during the day time.
\item (1998)  Explain briefly three different types of radio waves travelling from a transmitting station to a receiving antenna.
\item (2000)  Explain why Audio amplification is necessary for a practical radio set.
\item (2013)  An electron gun fires electrons at the screen of a TV tube. The electrons start from rest and are accelerated through a potential difference of $ 30$ kV. What is the speed of impact of electrons on the screen of the picture tube?
\item (2013)  Briefly explain why long distance radio broadcasts make use of short wave bands.
\item (2014)  Give the meaning of the following terms:
 \begin{itemize}
\item Bandwidth
\item  Amplitude modulated carrier wave
\end{itemize}
\item (2014)  Sketch the frequency spectrum for $ 1500$ m radio waves modulated by $ 4$ kHz audio signal.
\item (2014)  List down two advantages of digital signals over analogue signals.
\item (2014)  A carrier of frequency $ 800$ kHz is amplitude modulated by frequencies ranging from $ 1$ kHz to $ 10$ kHz.  What frequency range does each sideband cover?
\item (2015)  Give one advantage of frequency modulation (FM) as compared to amplitude modulation ( AMT).
\item (2015)  Briefly explain the importance of bandwidth of an amplitude modulation (AM) signal.
\item (2015)  State the function of a modulator in radios.
\item (2015)  Sketch a block diagram to show the general plan of any communication system.
\item (2015)  The amplitude modulated (AM) broadcast band ranges from $ 450$ to $ 1200$ kHz. If each station modulates with audio frequencies up to $ 5.5$ kHz, determine the
 \begin{itemize}
\item  Bandwidth needed for each station.
\item  Total bandwidth available. 
\end{itemize}
\item (2017)  List three basic elements of communication system. 
\item (2018)  Identify two difficulties which would arise when two straight wires are used to transmit electricity direct from the source to the city station. 
\item (2019)  Identify three basic elements of a communication system. 
\item (2019)  Why sky waves are not used for transmission of TV signals? 
\end{itemize}

\end{document}