
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{5.3 - Thermal Convection}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (2000)  Write down a formula for the rate of cooling under natural convection and define all the symbols used. 
\item (2007)  State Newton’s law of cooling and give one limitation of the law.
\item (2007)  A body initially at $ 70^{\circ}$C cools to a temperature of $ 55^{\circ}$C in $ 5$ minutes. What will be its temperature after $ 10$ minutes given that the surrounding temperature is $ 31^{\circ}$C ? (Assume Newton’s law of cooling holds true)
\item (2010)  Define thermal convection.
\item (2010)  State Newton’s law of cooling.
\item (2010)  A glass disc of radius $ 5$ cm and uniform thickness of $ 2$ mm had one of its sides maintained at $ 100^{\circ}$C while copper block in good thermal contact with this side was found to be $ 70^{\circ}$C . The copper block weighs $ 0.75$ kg. The cooling of copper was studied over a range of temperature and the rate of cooling at $ 70^{\circ}$C was found to be $ 16.5$ K$/$min. Determine the thermal conductivity of glass.
\item (2013)  A person sitting on a bench on a calm hot summer day is aware of a cool breeze blowing from the sea. Briefly explain why there is a natural convection?
\item (2013)  A cup of tea kept in a room with temperature of $ 22^{\circ}$C cools from $ 66^{\circ}$C to $ 63^{\circ}$C in $ 1$ minute. How long will the same cup of tea take to cool from the temperature of $ 43^{\circ}$C to $ 40^{\circ}$C under the same condition?
\item (2014)  Define thermal convection.
\item (2014)  Prove that at a very small temperature difference, $ \Delta T=T_{b}$ – $ T_{s}$ ,  Newton's law of cooling obeys the Stefan’s law, whereby $ T_{b}$ , is the temperature of the body and $ T_{s}$ is the temperature of the surrounding. 
\item (2016)  Briefly explain why forced convection is necessary for excess temperate less than $ 20$ K? 
\item (2016)  State Newton’s law of cooling. 
\item (2016)  A body cools from $ 70^{\circ}$C to $ 40^{\circ}$C in $ 5$ minutes. If the temperature of the surroundings is $ 10^{\circ}$C , Calculate the time it takes to cool from $ 50^{\circ}$C to $ 20^{\circ}$C.  
\end{itemize}

\end{document}