
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{\textbf{5.4 - Thermal Radiation}}
\author{PJ Gibson - Peace Corps Tanzania}
\date{May 2020}

\begin{document}

\maketitle

\begin{itemize}
\item (2007)  What is blackbody radiation of a given body?
\item (2007)  Explain why heat may just mean infrared.
\item (2007)  State Prvost's theory of heat exchange.
\item (2007)  What is Wien's displacement law?
\item (2007)  The sun's surface temperature is about $ 6000$ K.  The sun's radiation is maximum at wavelength of $ 0.5\times10^{-6}m$ .  A certain light bulb filament emits radiation with maximum wavelength of $ 2\times10^{-6}m$ .  If both the surface of the sun and of the filament have the same emissive characteristics, what is the temperature of the filament?
\item (2010)  State Stefan’s law of thermal radiation.
\item (2010)  A solid copper sphere cools at the rate of $ 2.8^{\circ}$C$/$min when its temperature is $ 127^{\circ}$C. At what rate will a solid copper sphere of twice the radius cool when its temperature is $ 227^{\circ}$C? In both cases the surroundings are kept at $ 27^{\circ}$C and conditions are such that  Stefan’s law may be applied.
\item (2010)  Explain the observation that a piece of wire when steadily heated up appears reddish in color before turning bluish. 
\item (2013)  A black body of temperature $ \theta $ is placed in a blackened enclosure maintained at a temperature of $ 100^{\circ}$C. When its temperature rises to $ 30^{\circ}$C the net rate of loss of energy from the body was found to be $ 10$ Watts. Find the power generated by the body at $ 50^{\circ}$C if the energy exchange takes place solely by the process of forced convection.
\item (2013)  Write down three laws governing the black body radiation.
\item (2015)  The element of an electric fire with an output of $ 1000$ W is a cylinder of $ 250$ mm long and $ 15$ mm in diameter. If it behaves as a black body, estimate its temperature.
\item (2016)  Briefly explain why: 
 \begin{itemize}
\item A body with large reflectivity is a poor emitter. 
\item The earth without its atmosphere would be too cold to live.
\end{itemize}
\item (2016)  What is meant by thermal radiation?
\item (2016)  Why is the energy of thermal radiation less than that of visible light? 
\item (2016)  A body with a surface area of $ 5.0 $ cm$ ^{2}$ and a temperature of $ 727^{\circ}$C radiates $ 300$ joules of energy in one minute. Calculate its emissivity.
\item (2017)  State the following according to heat exchange:
 \begin{itemize}
\item  Prevost’s theory. 
\item Wien's displacement law.
\end{itemize}
\item (2018)  Why during emission of radiations from black body its temperature does not  reach zero Kelvin? 
\item (2018)  A black ball of radius $ 1$ m is maintained at a temperature of $ 30^{\circ}$C . How much  heat is radiated by the ball in $ 4$ seconds? 
\item (2019)  Sketch the graph to illustrates how the energy radiated by a black body is distributed among various wavelengths. 
 \begin{itemize}
\item What information would be drawn from the graph above? Give three points.
\end{itemize}
\item (2019)  At what temperature will the filament of a $ 10$ W lamp operate if it is supposed to be a perfectly black body of area  $ 1 $ cm$ ^{2}$ ? 
\end{itemize}

\end{document}