
\documentclass{article}
\usepackage[a4paper, total={6in, 8in}]{geometry}
% \usepackage[utf8]{inputenc}
\usepackage{abstract}
\title{THE UNITED REPUBLIC OF TANZANIA

NATIONAL EXAMINATIONS COUNCIL

ADVANCED CERTIFICATE OF SECONDARY EDUCATION EXAMINATION

\textbf{2016 PHYSICS 2}}
\author{Transcribed by:  PJ Gibson}

\begin{document}

\maketitle

\begin{itemize}
\item Distinguish between static pressure, dynamic pressure and total pressure when applied to streamline or laminar fluid flow and write down expression at a point in the fluid in terms of the fluid velocity v, the fluid density $ \rho $ , pressure $ P$ and the height $ h$ , of the point with respect to a datum.  
\item The static pressure in a horizontal pipeline is $ 4.3 \times 10^{4}$ Pa, the total pressure is $ 4.7 \times 10^{4}$ Pa and the area of cross-section is $ 20 $ cm$ ^{2}$ . The fluid may be considered to be incompressible and non-viscous and has a density of $ 1000$ kg$/$m$ ^{3}$ .  Calculate the flow velocity and the volume flow rate in the pipeline.
\item State Newton’s law of viscosity and hence deduce the dimensions of the coefficient of viscosity.
\item In an experiment to determine the coefficient of viscosity of motor oil, the following measurements are made:
 \begin{itemize}
\item Mass of glass sphere $ =1.2 \times 10^{-4}$ kg.
\item Diameter of sphere $ =4.0 \times 10^{-3}$ m.
\item Terminal velocity of sphere $ =5.4 \times 10^{-5}$ m$/$s.
\item Density of oil $ =860$ kg$/$m$ ^{3}$
\item Calculate the coefficient of viscosity of the oil.
\end{itemize}
\item Briefly explain the carburetor of a car as applied to Bernoulli’s theorem.
\item Three capillaries of the same length but with internal radii $ 3R$ , $ 4R$ , and $ 5R$ are connected in series and a liquid flows through them under streamline conditions.  If the pressure across the third capillary is $ 8.1$ mm of liquid, find the pressure across the first capillary.
\item Give reasons for the following observations as applied in fluid dynamics.
 \begin{itemize}
\item A flag flutter when strong winds are blowing on a certain day.
\item A parachute is used while jumping from an airplane.
\item Hotter liquids flow faster than cold ones.
\end{itemize}
\item Define the following terms:
 \begin{itemize}
\item Intensity of sound
\item Beats
\item Ultrasonic
\item Overtones
\end{itemize}
\item A steel wire hangs vertically from a fixed point, supporting a weight of $ 80$ N as its lower end.  The length of the wire from the fixed point to the weight is $ 1.5$ m.  Calculate the fundamental frequency emitted by the wire when it is plucked if its diameter is $ 0.5$ mm. 
\item Give any two applications of ultrasonic as applied to sound waves.
\item Ultrasound of frequency $ 4.0$ MHz is incident at an angle of $ 30^{\circ}$ to a blood vessel of diameter $ 1.6$ mm.  If a Doppler shift of $ 3.2$ kHz is observed, calculate the blood flow velocity and the volume rate of blood flow.  Assume that the speed of ultrasound is $ 1.5$ km$/$s.
\item The absorption spectrum of a faint galaxy is measured and the wavelength of one of the lines identified as the calcium $ H$ line is found to be $ 478$ nm.  The same line has a wavelength of $ 397$ nm when measured in a laboratory. 
 \begin{itemize}
\item Is the galaxy moving towards or away from the observer on the Earth?
\item Determine the speed of the galaxy relative to observer on the Earth.
\end{itemize}
\item State the principle of:
 \begin{itemize}
\item Superposition of waves
\item Huygens construction of wave fronts.
\end{itemize}
\item The incident parallel light is a monochromatic beam of wavelength $ 450$ nm.  The two slits $ A$ and $ B$ have their centres, a distance of $ 0.3$ mm apart.  The screen is situated a distance of $ 2.0$ m from the slits.
 \begin{itemize}
\item Calculate the spacing between fringes observed on the screen.
\item How would you expect the pattern to change when the slits $ A$ and $ B$ are each made wider?
\end{itemize}
\item Describe the formation of interference patterns by using Newton’s rings experiment.
 \begin{itemize}
\item Calculate the radius of curvature of a Plano-convex lens used to produce Newton’s rings with a flat glass plate if the diameter of the tenth dark ring is $ 4.48$ mm, viewed by normally reflected light of wavelength $ 5.0 \times 10^{-7}$ m.  What is the diameter of the twentieth bright ring?
\end{itemize}
\item Define the following terms:
 \begin{itemize}
\item Free surface energy
\item Capillary action
\item Angle of contact
\end{itemize}
\item Briefly explain the following observations:
 \begin{itemize}
\item Soap solution is a better cleansing agent than ordinary water.
\item When a piece of chalk is put into water, it emits bubbles in all directions.
\end{itemize}
\item Two spherical soap bubbles are combined.  If v is the change in volume of the contained air, $ A$ is the change in total surface area, show that $ 3P_{A}V+4A T=0$ . Where $ T$ is the surface tension and $ P_{A}$ is the atmospheric pressure.
\item There is a soap bubble of radius $ 3.6 \times 10^{-4}$ m in air cylinder which is originally at a pressure of $ 10^{5}$ N$/$m$ ^{2}$ . The air in the cylinder is now compressed isothermally until the radius of the bubble is halved. Calculate the pressure of air in the cylinder.
\item What is strain energy?
\item State coulomb’s law of electrostatics.
\item Define electric field strength, $ E$ at any point.
\item Mention two common properties of electric field lines.
\item Two capacitors $ C_{1}$ and $ C_{2}$ each of area $ 36$ cm$ ^{2}$ separated by $ 4$ cm have capacities of $ 6$ $\mu$C and $ 8$ $\mu$C respectively.  The capacitor $ C_{1}$ is charged to a potential difference of $ 110$ V whereas the capacitor $ C_{2}$ is charged to a potential difference of $ 140$ V.  The capacitors are now joined with plates of like charges connected together.
 \begin{itemize}
\item What will be the loss of energy transferred to heat in the connecting wires?
\item What will be the loss of energy per unit volume transferred to heat in the connecting wires?
\end{itemize}
\item Define the following terms:
 \begin{itemize}
\item Capacitance
\item Charge density
\item Equipotential surface
\end{itemize}
\item By using the coulomb’s law of electrostatics, derive an expression for the electric field strength $ E$ , due to a point charge if the material is surrounded by a material of permittivity $ \epsilon $ , and hence show how it relates with charge density $ \rho $ .
\item Describe the structure and the mode of action of a simplified version of the Van de Graaff generator.
\item Identify any three factors on which the capacitance of parallel plate capacitor depends.
\item A proton of mass $ 16.7 \times 10^{-28}$ kg falls through a distance of $ 2.5$ cm in a uniform electric field of magnitude $ 2.65 \times 10^{4}$ V$/$m.  Determine the time of fall if the air resistance and the acceleration due to gravity, $ g$ , are neglected.
\item A parallel plate capacitor is made of a paper $ 40$ mm wider and $ 3.0 \times 10^{-2}$ mm thick.  Determine the length of the paper sheet required to construct a capacitance of $ 15$ $\mu$F , if its relative permitting is $ 2.5$ .
\item State any three magnetic components of the earth’s magnetic field.
\item The horizontal and vertical components of the Earth’s magnetic field at a certain location are $ 2.7 \times 10^{-5}$ T and $ 2.0 \times 10^{-5}$ T respectively.  Determine the Earth’s magnetic field at the location and its angle of inclination i.
\item State the following laws or theorems as applied in magnetism.
 \begin{itemize}
\item Biot-Savart law
\item Ampere’s theorem
\end{itemize}
\item Draw hysteresis loops diagram for soft iron and hard steel and use them to discuss permanent magnets.
\item Define permeability constant.
\item Derive an expression for the magnetic flux density $ B$ at the centre of the circular coil of radius $ r$ and $ N$ turns placed in air carrying a current i.
\item The diameter of a $ 40$ turn circular coil is $ 16$ cm and it has a current of $ 5$ A.  Calculate:
 \begin{itemize}
\item The magnetic induction at the centre of the coil
\item The magnetic moment of the coil.
\item The torque action on the coil if it is suspended in a uniform magnetic field of $ 0.76$ T such that its plane is parallel to the field.
\end{itemize}
\item Briefly explain the production of X-rays.
\item List down any three uses of X-rays.
\item How are the intensity and penetrating power of an X-ray beam controlled?
\item An X-ray tube, operated at a d.c. potential difference of $ 60$ kV , produces heat at the target at the rate of $ 840$ W .  Assuming $ 0.65\%$ of the energy of the incident electrons is converted into X-radiation, calculate:
 \begin{itemize}
\item The number of electrons per second striking the target.
\item The velocity of the incident electrons.
\item The energy of incident electrons
\end{itemize}
\item Show that the possible energy levels (in joules) for the hydrogen atom are given by the formula: $ E_{n}=-k^{2}(2\pi^{2}me^{4}/h^{2})(1/n^{2})$ .  Where $ m$ is the mass of electron, $ e$ is the electronic charge, $ h$ is the Planck’s constant, $ k=1/4\pi\epsilon _{0}$ and $ \epsilon _{0}$ is the permittivity constant of vacuum.  
 \begin{itemize}
\item What does the negative sign signify in the formula above?
\end{itemize}
\item The first member of the Balmer series of hydrogen spectrum has wavelength of $ 6563 \times 10^{-10}$ m. Calculate the wavelength of its second member.
\item Differentiate natural radioactivity from artificial radioactivity.
\item Name three applications of radioisotopes in medicine.
\item State two conditions for stability of nuclides referring to light nuclides and heavy nuclides.
\item Derive an expression for the half-life using the radioactive decay law.
\item What is carbon $ -14$ ?  Explain its production and how it is used in the dating process.
\item Living wood has an activity of $ 16.0$ counts per minute per gram of carbon.  A certain sample of dead wood is found to have an activity of $ 18.4$ counts per minute for $ 4.0$ grams.  Calculate the age of the sample of dead wood.  Assume the half-life of carbon $ -14$ is $ 5568$ years.
\end{itemize}

\end{document}